\documentclass[12pt]{article}
\usepackage{enumitem}
\usepackage[paperheight=11in,paperwidth=8.5in,margin=0.75in]{geometry}

\usepackage{fancyhdr}
\pagestyle{fancy}
\rhead{Shrinu Kushagra (NSERC PIN: 504441)}

\setlength\parindent{0pt}
\begin{document}
%\linespread{0.901}
%\hspace{0.5cm} 
\begin{itemize}[leftmargin=*]
\item[Q] \textbf{What progress was made during the previous year toward completing the research program or degree requirements (courses, comprehensive examination, thesis, etc.)? Did this progress meet or surpass the objectives set at the beginning of the year? Explain.}
\item[A] \vspace{-0.1in} This year all the requirements towards the degree completion were except for the PhD thesis defence. The date of the phd defence  has been set to $29^{th}$ May, 2019. All the formalities and paperwork including finding an external examiner, an internal-external examiner, submitting a thesis draft to all the members of the committee have been completed. The progress has met the expectations set for last year which was to finish my phd by the end of summer 2019.

\item[Q] \textbf{What progress was achieved during the previous year with respect to professional development (conference presentations, publications, etc.)?}
\item[A] With respect to professional development, good progress has been made. We had two papers accepted at two conferences this year. The first one was at ICDE (International Conference on Data Engineering) in Hong Kong. ICDE is a one of the top three conferences in the field of databases. Our result on applying clustering theory for the problem of data deduplication was accepted for a full-presentation at the conference. The typical acceptance for this conference is $18-25\%$. Another paper pursuing the same line of work but from a theoretical perspective was accepted at AISTATS (22nd International Conference on Artificial Intelligence and Statistics), Okinawa, Japan. This is again one of the top three conferences in the field of Machine Learning with an acceptance rate in the range $25-30\%$. 

We have another paper on `Clustering in the presence of noise'. At the time of writing, we are preparing to submit this work to the Journal of Machine Learning Research. 

\item[Q] \textbf{What degree requirements (courses, comprehensive examination, thesis, etc.) still need to be completed? Please specify the deadlines for completion, as well as specific objectives, for the coming year.}
\item[A] As mentioned in the answer of the first question, the defense of thesis is the final degree requirement that is yet to be met. I will defend it on May $29^{th}$, 2019 and finish all the degree requirements by the end of the coming summer, 2019. 

\item[Q] \textbf{Other comments (if any).}
\item[A] I am planning to have a startup (after I finish my degree requirements) which applies my learning and research in the field of machine learning over the last several years to try to solve a real-world problem in the fitness and sports industry. Ideally, I would love to start this in the Toronto-Waterloo region and not have to move to the US. The Govt. of Canada  has generally been very welcoming and encouraging of all the professional endeavours. Since, the time I first set foot on canadian soil ($1^{st}$ Sept., 2012), the canadian people have always treated me and other immigrants with love, respect and dignity. The startup is my way to try to pay back that gratitude towards the people, in a small way. Any help in this regard would be greatly appreciated.
\end{itemize}
\end{document}