\documentclass[12pt]{article}
\usepackage{enumitem}
\usepackage[paperheight=11in,paperwidth=8.5in,margin=0.75in]{geometry}

\usepackage{fancyhdr}
\pagestyle{fancy}
\rhead{Shrinu Kushagra (NSERC PIN: 504441)}

\setlength\parindent{0pt}
\begin{document}
%\linespread{0.901}
%\hspace{0.5cm} 
\begin{itemize}[leftmargin=*]
\item[Q] \textbf{What progress was made during the previous year toward completing the research program or degree requirements (courses, comprehensive examination, thesis, etc.)? Did this progress meet or surpass the objectives set at the beginning of the year? Explain.}
\item[A] \vspace{-0.1in} The courses and comprehensive examination were already completed before the beginning of the previous year. Towards thesis completion, the following goals were set. My research is in the theory of clustering. We had developed a framework for incorporating user expertise into the problem of clustering. Our goal at the start of the year was to extend our original framework to handle more practical situations. Our progress definitely met this goal. So far, we have been able to extend and apply our original framework on the problem of 'data de-duplication' which is a very important problem for the people in the database community.  

\item[Q] \textbf{What progress was achieved during the previous year with respect to professional development (conference presentations, publications, etc.)?}
\item[A] As mentioned above, we have extended our framework to the problem of data de-duplication. We have prepared a manuscript which we hope to be able to submit to one of the top machine learning conferences within the next month. We already submitted another research paper to ICML (one of the top tier conferences in ML). This work tackles the problem of clustering in the presence of noisy points. Again, this paper relates to our original goal, ``{\em extend the clustering framework to handle more realistic situations}".  

\item[Q] \textbf{What degree requirements (courses, comprehensive examination, thesis, etc.) still need to be completed? Please specify the deadlines for completion, as well as specific objectives, for the coming year.}
\item[A] Two degree requirements are still left. The first is the {\em Seminar} requirement. PhD students are required to give three seminars on topics related to their research. So far, I have given two seminar talks. I expect to give the third and final one by the end of June 2018. The second is the {\em Thesis} requirement. I expect that after finishing the `clustering for data de-duplication' project, I would have accomplished the results required to be able to defend my thesis. Over the last year, I with the help of my advisor have laid down the theoretical framework/work required to be able to accomplish this task. I expect that within the next six months (end of August 2018) we will be able to drive this project to its conclusion. I would then spend the next 4-5 months writing my thesis and defending my thesis in Winter 2019 (sometime in Feb 2019). 

\item[Q] \textbf{Other comments (if any).}
\item[A] I have also been a part of the Squash Team at the University of Waterloo winning the silver medal at the Ontario University Athletics Squash Championships. I would keep playing for my school this year too.   
\end{itemize}
\end{document}